### 2016-05-04 paper/intro.tex

\section{Introduction}
\input{fig_regex-definition}
\input{fig_lens-syntax}
\input{fig_lens-typing}
\input{fig_lens-semantics}
We have regular expressions as normal regular expressions.
We expand it with having user defined data types as well.
We have lenses, defined in Figure~\ref{fig:lens-syntax}, typed as in Figure~\ref{fig:lens-typing}.
These lenses have underlying functions, given via Figure~\ref{fig:lens-semantics}.
We would like to be able to synthesize these lenses automatically, given a
specification as two regular expressions, an a set of values that are
mapped to each other.