### 2016-05-29 paper/fig_dnf-regex-syntax.tex
added latexes

\begin{figure}
\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}

% DNF_REGEX
(Atoms)& \Atom{} & \GEq{} & \RegexVariable{} & variable\\
& & &\GBar{} \DNFRegex{}* & iterate \\
(Conjunct)& \Conjunct{} & \GEq{} &
$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & conjunct\\
(DNF Regex)& \DNFRegex{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & dnf regex\\
\end{tabular}
\caption{DNF Regex Syntax}
\label{fig:dnf-regex-syntax}
\end{figure}