### 2016-05-31 paper/fig_dnf-regex-syntax.tex
added in some clenaer stuff

\begin{figure}
\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}

% DNF_REGEX
(Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \DNFRegex{}* & Iterate DNF\\
(Conjunct)& \Conjunct{},\ConjunctAlt{} & \GEq{} &
$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & Conjoin\\
(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & DNF Or\\
\end{tabular}
\caption{DNF Regex Syntax}
\label{fig:dnf-regex-syntax}
\end{figure}