### 2016-06-07 paper/evaluation.tex
added some latex bug fixes

\section{Evaluation}

\input{fig_runtimes-table}
\subsection{Speed}
We ran

\input{fig_examples-required-table}
\subsection{Importance of Examples}

\input{fig_expanded-time-table}
\subsection{Importance of User Defined Data Types}

Evaluation~\ref{fig:examples-required-table}~\ref{fig:runtimes-table}~\ref{fig:expanded-time-table}.