### 2016-06-07 paper/fig_examples-required-table.tex

\begin{figure}
\centering
%\begin{tabular}{|l|c|}

\csvreader[tabular=|l|l|,
    table head=\hline\textbf{Test} & \textbf{Example Count} \\\hline,
    late after line = \\\hline]%
{generated-data/data.csv}{Test=\Test,ExamplesRequired=\ExamplesRequired}%
{\Test{} & \ExamplesRequired{}}%

%\hline\textbf{Test} & \textbf{Examples Required}
%\\\hline\hline
%\csvreader[head to column names]{generated-data/data.csv}{}
%{\Test& \ExamplesRequired\\\hline}
%\end{tabular}
\caption{Examples Required Table}
\label{fig:examples-required-table}
\end{figure}