### 2016-06-09 paper/introduction.tex

\bcp{May I reiterate my request to put all the text and figures into one big
.tex file?  Makes finding things, global replace, renaming sections, etc. very much easier.  :-)}

\section{Introduction}