### 2016-06-10 paper/fig_runtimes-table.tex

\begin{figure}
\centering
%\begin{tabular}{|l|c|}

\csvreader[tabular=|l|l|,
    table head=\hline\textbf{Test} & \textbf{Time (s)} \\\hline,
    late after line = \\\hline]
{generated-data/data.csv}{Test=\Test,ComputationTime=\ComputationTime}
{\Test{} & \ComputationTime{}}

%\hline\textbf{Test} & \textbf{Examples Required}
%\\\hline\hline
%\csvreader[head to column names]{generated-data/data.csv}{}
%{\Test& \ExamplesRequired\\\hline}
%\end{tabular}
\caption{Runtime Table\bcp{Hard to read, with all the zeros.  Maybe better
    to display as a (log-scale) graph?  More seriously, we'll need to
    explain the examples someplace, say why we chose them, maybe group them
    in some way, ...  Why ws this the right set of tests?}}
\label{fig:runtimes-table}
\end{figure}