### 2016-07-03 paper/extended-example.tex
edits to synthesis example

\section{Extended Example}

Consider building a lens which converts between between citations given
in EndNote format, and in BibTex format.  For simplicity, we are only
considering the author and title fields.  The synthesis problem for these
two regular expressions is provided below.

\begin{figure}
\begin{lstlisting}
Name    = [A-Z][a-z]*
Initial = [A-Z] "."
Word    = ([A-Z] | [a-z] | "-" | "," | ":")+
Title   = Word (" " Word)*

FN      = Name", " Name (" " (Name | Initial))*
Type    = "@inproceedings" | "@article" 
BibTex  = 
  Type "{" 
    Word ","
    ("" | ("title={" Title "},"))
    ("" | ("author={" FN (" and " FN)* "},")) 
  "}"

EName   = Name (" " (Name | Initial))* (" " Name)
Class   = "Conference Paper" | "Journal Article"
EndNote = "%0 " Class "\n"
	 ("%A " EName "\n")*
	 ("" | "%T TITLE")
	  "%1 " Word "\n"

BibEndConv : BibTex <=> EndNote = ?
\end{lstlisting}
\caption{Synthesis Problem: Simplified Bibtex to EndNote}
\end{figure}

We are trying to synthesize \BibEndConv{}.