### 2016-07-03 paper/implementation.tex

\section{Implementation}

\subsection{Synthesis Algorithm}
\begin{lstlisting}[mathescape]

let gen_dnf_lens (q:priority_queue) (es:examples)
    : dnf_lens$_\bot$ =
  let (dnf_r,dnf_s,q) = q.pop()
  let dnf_lens_option =
    gen_tds_dnf_lens dnf_r dnf_s es in
  begin match

let gen_lens (r:regex) (s:regex) (es:examples) : lens$_\bot$ =
  let dnf_r = $\ToDNFRegex$(r) in
  let dnf_s = $\ToDNFRegex$(s) in
  let dnf_lens = gen_dnf_lens(dnf_r, dnf_s, es) in
  to_lens(dnf_lens)
\end{lstlisting}