### 2016-09-12 paper/conclusion.tex
added variable information on master file to each file

\section{Conclusion}

Conclusion

%%% Local Variables:
%%% TeX-master: "main"
%%% End: