### 2016-09-12 paper/discussion.tex

\section{Discussion}

Discussion

%%% Local Variables:
%%% TeX-master: "main"
%%% End: