### 2016-09-16 paper/conclusion.tex
added rough draft of all sections

\section{Conclusion}

We have coded up an implementation of type directed synthesis on a domain
specific language of bijective lenses.  This program allows users to input two
regular expressions, and a few input/output examples, and synthesizes a
bijective lens between those regular expressions, matching those input/output
examples.  While lenses are very useful for their guarantees, they are a
difficult domain specific language to code in, in practice.  This program
provides an easier interaction model for the creation of lenses through using
synthesis as the way to interact with the programmer.

This synthesis algorithm is competitive with state of the art tools for
synthesis on strings transformations.  This gain is made through requiring a
larger amount of user interaction with the synthesis algorithm.  However, in the
for this increased interaction, the program provides the capabilities of
synthesizing very complicated transformations, and provides stronger guarantees
about correctness on inputs dissimilar to those provided in the examples.

%%% Local Variables:
%%% TeX-master: "main"
%%% End: