### 2016-09-22 paper/introduction.tex

\section{Introduction}

The growing dependence on big data makes tools for reliably parsing,
printing, cleaning, and transforming data increasing important.
One such class of tools are \emph{lenses}~\cite{Focal2005-long}, bi-directional
programs that help users solve the classic ``view update problem.''
More specifically, a lens provides a user with two functions, a
\emph{get} function and a \emph{put} function.  The \emph{get}
function translates native data into a new format, or \emph{view}, for
the user.  If the user updates the new view, the corresponding
\emph{put} function can translate the edited data back into
the native format.

Bidirectional programming is used heavily in practice.
Augeas~\cite{?}, a configuration editing system for linux, uses lens combinators
to provide editable views of configuration files.
Bixid~\cite{?} is a language developed for expressing bidirectional XML
transformations.
XSugar~\cite{?} is used to convert bidirectionally between XML and non-XML data,
and is used to convert between EpiDoc~\cite{?} and Leiden+~\cite{?} formats for
encoding ancient documents~\cite{?}.
Serializers and deserializers combine to make a bidirectional program, though
they are rarely written in a domain specific language for bidirectional
programming.

Lenses are popular because they typically come with strong guarantees
about the get-put (and put-get) round-trip behavior.  Such guarantees
help users guard against data corruption while reading and editing
high-value data sources.  These guarantees are often achieved through
the use of domain-specific languages with strong type systems that
guarantee invertability of lens programs~\cite{?,?,?}.  Unfortunately, such
lens languages are currently quite non-standard.  The
process of writing a lens is challenging, as a user must think
about data conversion in two directions at once --- an
unusual task for the everyday programmer.  As a consequence, 
for the casual user of lenses, learning the language, its type system and its
unorthodox programming style is time-consuming and, in the end,
a disincentive to adopting the technology, despite its advantages.

In order to overcome the steep learning curve involved in programming
with lenses, we have developed a new system for synthesis of lenses
from their types and a collection of representative examples.  More
specifically, we focus on algorithms for the synthesis of bijective
string lenses.  The types for such lenses are regular expressions,
which describe the format of the source and target strings.  Given
such types, and a very small number of representative examples (sometimes
no examples at all), we synthesize a well-typed lens that implements
a bijection between data formats.

Recently, researchers have developed a number of new algorithms for
type-directed synthesis~\cite{?,?,?}.  Many of those
algorithms~\cite{osera+:pldi15,frankle+:popl16,armando+:pldi16} are
directed in part by the syntax of types.  For instance, to find an
expression with type $A \rightarrow B$, such algorithms will assume
that the expression in question is a function with the form $\lambda
x{:}A. e$, and they will continue the search for an expression $e$
with type $B$ under the hypothesis that the parameter $x$ has type
$A$.  In the case of the simply-typed lambda calculus, such an
approach reduces the complexity of the search for terms inhabiting the
type without losing completeness.  Lenses are defined by a richer type system
than the simply typed lambda calculus,
with their regular expression types, as
there are many equivalent ways to express the same regular expression.
Unfortunately, this makes simple
syntax-directed algorithms not directly applicable.

\paragraph*{The Central Challenge.}
To illustrate the central challenge facing the design of our synthesis system
in greater detail,
consider the concatenation lens, $\Lens_1 \Concat \Lens_2$,
one of the simplest string lens combinators.  
When viewed from left to right, this lens defines a function \emph{get}, which,
given a string 
$\String$, divides $\String$ into two substrings $\String_1$ 
and $\String_2$, applies $\Lens_1$ to the first, yielding $t_1$,
and $\Lens_2$ to the second, yielding $t_2$.  When viewed from
right to left, it defines a second function \emph{put}, which maps
data back from strings $t$ to strings $s$ in an analogous way.

The type system for lenses ensures such \emph{get} and \emph{put} functions
satisfy strong invertibility properties.  In our case, the two functions
form a bijection.  In general, to 
describe the type of a lens, we write $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, which may be read ``lens $\Lens$ converts back and forth between
data in the language of regular expression $\Regex$ and data in the language
of regular expression $\RegexAlt$''.  More specifically, to give a type
to the concatenation lens, we use the following typing rule.
\[
\inferrule
{
\Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
\Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
\UnambigConcatOf{\Regex_1}{\Regex_2}\\
\UnambigConcatOf{\RegexAlt_1}{\RegexAlt_2}
}
{
\ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1\Regex_2 \Leftrightarrow \RegexAlt_1\RegexAlt_2
}
\]
Above, the notation $\UnambigConcatOf{\Regex_1}{\Regex_2}$ asserts that $\Regex_1$ and $\Regex_2$ are
unambiguously concatenable, meaning there is
exactly one way to divide up any
string $\String$ in the language of $\Regex_1\Regex_2$ into $\String_1$
and $\String_2$ such that $\String_1$ is in the language of $\Regex_1$
and $\String_2$ is in the language of $\Regex_2$.  These unambiguous
concatenation constraints guarantee the \emph{get} and \emph{put}
functions are indeed functions and that they form a bijection if
the underlying lenses are bijective.

Now, with these typing rules in hand, consider trying to search for
a lens with the type $(A B) C \Leftrightarrow A' (B' C')$ where
we know of lenses with types 
$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
$C \Leftrightarrow C'$.  A straightforward,
syntax-directed search suggests applying the rule for concatenation,
and looking for two simpler lenses with the types $A B \Leftrightarrow A'$
and $C \Leftrightarrow B' C'$ respectively.  Unfortunately, at this point,
we are simply stuck---we have no lenses with those types.  Of course,
the natural
solution to this problem
is to introduce a type conversion rule to the system:
\[
\inferrule
{
\Lens \OfType \Regex_1 \Leftrightarrow \Regex_2\\
\Regex_1 \equiv \Regex_1'\\
\Regex_2 \equiv \Regex_2'
}
{
\Lens \OfType \Regex_1' \Leftrightarrow \Regex_2'
}
\]
With such a type conversion rule, we could first 
convert the type $(A B) C \Leftrightarrow A' (B' C')$
to the better aligned $(A B) C \Leftrightarrow (A' B') C'$.
When we then proceed with structure decomposition using the
concatenation rule, we generate sub-problems involving
finding lenses for types
$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
$C \Leftrightarrow C'$, which is just what we had in hand.

Because equivalence of regular expressions is decideable, when
given a lens $\Lens$ with type $\Regex_1 \Leftrightarrow \Regex_2$,
it is easy to \emph{check} whether that lens also has type 
$\Regex_1' \Leftrightarrow \Regex_2'$, and lens programming languages
such as Boomerang \cite{foster:thesis} do so.  However, 
when we turn to synthesis and are given the lens type 
$\Regex_1' \Leftrightarrow \Regex_2'$ and the option to apply this
rule, the search space becomes unmanageably large --- we must guess
some arbitrary new regular expressions $\Regex_1$ and $\Regex_2$ that are
equivalent to  $\Regex_1'$ and $\Regex_2'$ respectively
and there are infinitely many such regular expressions to choose
from.

Furthermore, the language of lenses without a composition operator is not closed
under composition.  There exists lenses $\Lens_1 \OfType \Regex_1 \Leftrightarrow
\Regex_2$, $\Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3$ where there is no
lens $\Lens \OfType \Regex_1 \Leftrightarrow \Regex_3$ such that
$\SemanticsOf{\ComposeLensOf{\Lens_2}{\Lens_1}}=\SemanticsOf{\Lens}$.  So, if we
would like to be able to synthesize $\ComposeLensOf{\Lens_2}{\Lens_1} \OfType
\Regex_1 \Leftrightarrow \Regex_3$ then we
must also search through possible $\Regex_2$ for creating compositions.

In essence,
we must solve the problem of search \emph{in three dimensions}:  In one
dimension, we must search for a regular expression \emph{type}
with an appropriate ``syntactic shape''.  In a second dimension,
we must search for a lens \emph{expression} that has the given type and
is consistent with the user's examples.  In a final dimension, we must search
for intermediary regular expressions for composing lenses.

\paragraph*{Our Solution.}  Our solution to this problem is to define
a new language of lenses that
cuts down the synthesis search space \emph{in all dimensions}.
This language is designed so that its regular expression types are limited 
to those in a pseudo-canonical form.  This pseudo-canonical form
demands that regular expressions be defined as a flattened 
union of concatenated strings and atoms, where each atom is a user defined data
type or the Kleene star of an expression that is recursively in normal form.
When converted into this pseudo-canonical form, two regular expressions that 
differed only by associativity of concatenation will be syntactically equal.
For example, rather than searching for a lens with type 
 $(A B) C \Leftrightarrow A' (B' C')$, our algorithm will always search
for one with type  $(A B) C \Leftrightarrow (A' B') C'$

More generally, many, but not all, structural incongruities that
prevent the existence of lenses disappear.
Not all structural incogruities disappear because our regular expressions
are only \emph{pseudo-canonical}.  For instance, our pseudo-normal form
does not automatically unify lenses such as
$A B* \Leftrightarrow A (\epsilon + BB*)$ that result from an
unrolling of Kleene Star, for instance.
Consequently, we structure our synthesis algorithm in two parts:  The
first is a syntax-directed search for a lens based on the structure of the given
regular expression types and, if that fails, the second is a search,
using Conway's axioms~\cite{conway} for 
equivalent regular expression types that might allow a new search
in the first phase of the search to succeed.

Furthermore, in this new language, we can express transformations which were
previously inexpressible without composition.  It is still an open problem
whether this new language is closed under composition.  However, with the
increased expressibility, we have found searching for a central regular
expression to compose with to be unnecessary in practice.

The types
describing data formats may be very large --- they are regular
grammars describing complex data formats.  This is another complication that
arises in this synthesis setting, but not elsewhere in the literature on type
directed synthesis.  On the one hand, this is
very useful as such types provide a wealth of information that constrains
the search process effectively.  On the other hand, there is a lot of
data that must be managed carefully by the synthesis system.
One important observation is that users naturally construct
such types compositionally, using named abbreviations for subcomponents.
The structure inherent in these named abbreviations can be 
exploited by the synthesis algorithm to speed up the search dramatically.

For example, if the name $A$ appears once on both the left and the right,
the system will hypothesize that the identity lens can be used to convert
between occurrences of $A$.  On the other hand, if the name $A$ appears
on the left, but not on the right, the system will determine that the
name $A$ should be replaced by its definition.  Likewise, differing
numbers of occurrences of names can serve as a guide for unrolling
Kleene star or applying other regular expression equivalence axioms.
Hence, in general, to guide the transformation of regular expression
types, we define a pseudometric based on the occurrence of names.

\paragraph*{Contributions.}  We have designed and implemented
a new system for the synthesis of bijective string lenses.
In addition, this paper makes the following contributions:

\begin{itemize}
\item It defines a new language of lenses designed specifically for synthesis
with types based on pseudo-canonical regular expressions, with a term 
language to match.
\item We formalize the semantics of the language and prove it sound with
respect to an existing, standard language for bijective string lenses.
\item We designed a type-directed, two-dimensional 
algorithm for synthesizing lenses in our new domain-specific language.
\item We developed a collection of optimizations that speed up the basic
synthesis algorithm.
\item We have implemented the language and evaluated it on a range of
over 40 benchmarks including micro-benchmarks, a few examples from FlashFill,
and a number of examples from the Augeas system for transforming linux
file formats.  Our synthesis algorithm succeeds on all of our
benchmarks.  When it succeeds, synthesis time is typically less than 1 second.
\end{itemize}

%%% Local Variables:
%%% TeX-master: "main"
%%% End: