### 2017-09-11 papers/BijectiveLensSynth/popl_rebuttal_strategy_notes.txt
add in some popl rebuttal notes

Summarize points they all raise, and questions for rebuttal:

Quotient lenses allow for removing whitespace

Reviewer 1:
Specify and talk about QREs, transformations can be factored into 3 parts,
and we have a way of specifying 2 of those parts.  1 part that transforms
data into a canonical form, transformations into canonical forms and then
same transforms as usual in the middle.
We've experimented with this and our current inference algorithm works
off the shelf with it.

Rebuttal to reviewer B issues is that:
Are there any xforms that we can synth and flashfill can't
Start with that..., talk about nested iterations etc.
One of largest strengths is that it can synthesize really complex ones that
can't be synthesized by FF

Point to thing thing in Excel, as stated it can't synthesize nested stars

The DSL allows for nested stars, but it doesn't seem to work

Copy & paste text of their questions,
Differences between FF and Optician
How do we transform the examples, describe changes
Any chance of automating this or handling, cut & paste question

	2 major classes of changes we had to make: 1 is whitespace, other is
 	  projection of data
	Bijections have to preserve the amount of whitespace, this can be
	  automated by putting a wrapper around the xform that normalizes
	  the whitespace
	Our current system does that (this is called a quotient lens)
	Second transformation is forgetting information, and then restoring
	  after a round-trip.  We again believe the core of our algorithm
	  can be reused for this, and we are working on this still, but it's
	  future work.  Say how we modified the examples - added in the
	  extraneous data.  Again, a wrapper that deleted that and restored
	  it on the reverse transformation could handle it and is not hard
	  to write - believe can be automated, but need to do the research.

Reviewers are worried we're cheating about experiments

	Try and avoid subjective wording
	Concrete facts about what's going on
	In whitespace, we're making sure that each side has same amount of
	  whitespace