### 2018-09-19 papers/SymmetricLenses/response_popl19.md
do some clean

BCP: The first section is too long -- 500 words max.

Thank you for your comments!

We begin with brief responses to the most significant concerns, followed by a
(long, optional) appendix addressing individual questions point by point.

                                -------------

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

Beyond the format specifications and examples, we do not have specifications
that define the conversion functions. Hence, we do not have an automated way to
measure the correctness of our lenses (beyond checking that each synthesized
lens does indeed transform the given examples correctly). In this way, our work
is similar to other published synthesis-from-examples research, such as
Gulwani's work on FlashFill and related projects.

To check the correctness of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code. This was certainly a nontrivial task, as
    these generated lenses are quite large, but perhaps easier than it may
    initially sound. Some of the lenses' structure follows from the structure of
    the regular expressions that describe source and target format --- this is
    relatively easy to validate. However, it can be more difficult to check the
    interactions of the complex combinators: Are the swaps in the right place?
    Are the correct components disconnected? Are lenses merged in the correct
    way? These complex combinators often occupied a few dozen lines of code
    rather than a few hundred, making our task easier. (Certainly, checking the
    code is easier than creating it de novo.)
2.  We built a few unit tests to validate the more complex lenses. 
3.  We ran the lenses on a number of strings, taken from the unit tests and the
    examples, but with minor alterations. We did not add these tests as
    additional unit tests, but they served as additional validation for
    ourselves as we were constructing the benchmark suite.

We recognize we did not go into significant enough detail on this in the
submitted version, and will gladly add in such detail on validation in the final
version.

                                -------------

      - There were a number of related questions about our information theoretic
        measure: Is it useful? Why not use a syntactic metric? Why does a
        fixed distribution work?  Will knowing the data distribution help?

Having a guiding principle behind these heuristics is, in itself, useful. Ad-hoc
syntactic metrics that are present in other systems (and we claim: are typically
building towards our underlying principle). Finding a simple underlying
principle (be as bijective as possible, which we formalize via information
theory) is better than a metric that simply applies only to our domain. We claim
that information theory is a general-purpose way to group all these metrics
where people give higher cost to their "less bijective" combinators (avoiding
constants, for example).

A fixed distribution works because, the core thing we are aiming for is "more
bijective." A fixed distribution is merely a way to not unfairly weight one
conjunct or disjunct over another in the presence of not knowing other
information. Even with our inferred distribution, we are still aiming towards
"more bijective".

Knowing the distribution will help in certain scenarios, and hinder in others.
With a learned distribution, our information theoretic measure is closer to the
ground-truth, permitting our learned lens to, in fact, be "more bijective."
However, we recognize that, while we believe "more bijective" is a solid guiding
principle, there are certain situations where the most bijective lens is not the
desired one: this is why we must further use examples for specifications. We
could certainly imagine situations where a learned distribution would find the
wrong lens (without additional examples), and synthesizing with the fixed
distribution would find the correct one, and vice-versa. However, we claim that,
as we typically want "more bijective," and a learned distribution gets closer to
a ground-truth on that, a learned distribution likely would perform better.

===========================================================================
===========================================================================
Detailed responses to individual questions in the reviews follow...

===========================================================================
Review #43A

                                -------------

     - I have no significant complaints about this paper apart from Section 3 
       not making it clear that stochastic regular expressions are not a
       contribution of the authors (albeit it is mentioned in passing on p2).

Yes, we will make this clearer on p. 3 too.

                                -------------

     - 3. p5 l236,

          Shouldn't the second argument to `disconnect` be `company`?

We should have the second type be "" (to be consistent with what we use in the
final lens)

BCP: ?? 
AFM: In hindsight, I think I will just delete this question - it's more
of a "finding-errata" question than an actual one

                                -------------

     - 4. p6, l288,

          If I understood correctly, this parituclar lens `disconnect(...)`
          picks the default values at random, but is still valid because it
          satisfies the required lens type. If so, please make it clearer.

This is correct -- we will clarify.

                                -------------

     - 9. p16, l777,

          Could you give concrete examples of situations where all sequences
          are involved in sequence lenses, but there are examples that could
          make other lenses more useful?

Within a DNF lens itself, all sequences are *required* to be involved with
another sequence lens: otherwise when that side of the DNF SRE is provided -
what disjunct on the other side should that lens be mapped to? Of course, there
are other ways the disjunct can be mapped: but these must happen at a different
level of the synthesis algorithm. For example - the entirety of that DNF regular
expression (which itself could be deeply nested within other DNF regular
expression components) could be projected. Furthermore, a specific sequence
could be mapped to another, but all the components of that sequence lens
actually may be disconnects. However, all sequences *must* be involved in a
sequence lens in a well-typed DNF lens.

BCP: I'm not sure I understand the question! 
AFM: I also now realize I didn't. After rereading the section, I think this
response is more appropriate.

                                -------------

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is as a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Furthermore, we look for instances where a regular
expression may be present in one format, but is only contained within a
closed regular expression in the other, and we expand the closed RE
containing the missing RE in this case. We then make this apply in slightly
more places by also taking into account how deeply nested within stars
certain REs occur.

BCP: What is "closed"?
AFM: The internals of the Regular Expression are not exposed to GreedySynth,
they are the analogue of User Defined regular expressions in the previous paper,
but with this closed/open approach, we aren't dependent on where users place
variables (also making our integration into Boomerang significantly less hacky).

Full details can be found in the prior work on Synthesizing Bijective Lenses
(page 16).

                                -------------

     - 11. p19, l887

           Surely the definition of `CONTINUE(pq)` should be using `max`? Otherwise
           I don't see how the definition makes sense, because using surely `min` 
           will just pick 0 for t>=0, because the distance and the logarithm are
           positive values.

Right, this should be max!

                                -------------

     - 12. p21, l1022

           Which preference metric are the t=0, t=25, t=-25 benchmarks using?

These benchmarks are using the full "SS" mode. The only difference between these
benchmarks and the full mode is that termination parameters are ignored, and
are instead required to use the 0, 25, and -25 parameters.

                                -------------

      - Second, from what I remember from the literature on (ordinary) symmetric
        lenses, they do not lend themselves very well to analysis in terms of 
        spans of asymmetric lenses. With simple symmetric lenses, do you think
        the situation is better now that you have omitted the complement part?

TODO: Learn Category Theory

                                -------------

       - Thirdly, do the authors expect to be able to extend the proposed 
         synthesis algorithm to handle any of the many variants of updates-based
         lenses, such as edit lenses, delta lenses, and update lenses?

Yes, we do have some hope that the algorithm can be extended to handle such
variants.

Delta lenses would require additional work in understanding "shape"
constructs.  Currently our language only permits basic iteration, whereas
shapes allow for different types of mergings. Allowing this would require
changes to the AtomSynth algorithm, in that we would we would need to
discover the correct shapes of the data, and align them in the correct ways.

Both edit and update lenses would require new forms of specifications, since
both permit updates that are not merely pushing data through, but instead
are applying specific elements of an edit monoid.  (Also we would remove
complements from edit lenses).

===========================================================================
Review #43B

     - The main weakness is in the evaluation part. There are many 
       unclear/controversial parts regarding the evaluation. For example, the 
       lack of description of the benchmark environment; the way they test 
       whether the synthesised lens is correct; the description of the tasks to
       be synthesised. See detailed comments below. (This defect does not
       prevent it from being a good paper.)

We responded to the more substantive points here (the benchmark suite and
judging correctness) "above the fold" at the top of this response.

All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
GB of 1600 MHz DDR3 running macOS High Sierra.

                                -------------

     - The first subsection of Related Work did not give any reference to
       the related work except for Foster et al. 2007, which makes the
       paper unfriendly and hard to verify. The appendix is also in a
       messy.

We apologize for the lack of citations and the messy appendix. For the first
subsection of Related Work, we got carried away in making formal claims, and did
not include informal comparisons to other lens languages. We will make sure to
make such comparisons and citations in the next version of the paper. We will
similarly improve the appendix.

                                -------------

     - L11: The reference Martin Hofmann et al. (2015) does not help since it
       is not yet published and cannot be obtained from Internet. Martin
       Hofmann et al. (2011) may be the choice for this time.

Thank you, you are correct, we will change this citation.

                                -------------

     - L442: H(s) should be 0. Otherwise the entropy of the SRE in L452 is 2.

You are correct.

                                -------------

     - L484 and L485:
       I cannot see the reason to relax the restriction regarding unambiguous
       REs. What will happen if primitive lens combinators for unambiguous
       REs are composed with the ones for ambiguous REs? Since complex lenses
       are composed from (ambiguous and unambiguous) primitive lenses, in
       practice the restriction should be always there.

Within our synthesis algorithm, you are, in general, correct. However,
Boomerang exists beyond Optician, and permits lenses like id("a"|"a"). In
our extension to Boomerang, we do not remove this freedom.

                                -------------

     - L557 and below: Why some of the entropy is a single value while some
       of them seems ranging over an interval? If the interval is used to
       handle `merge_left`, then the entropy of all combinators should be an
       interval. Currently only `id`, `disconnect`, and `merge_left` yield
       intervals.

This was us trying to make our equations more readable, but we did not
sufficiently explain what our syntax meant. The calls to H(...) return pairs,
and when multiplied by constants, they act as follows:

  a x (b,c) = (a x b,a x c)
  (a,b) + (c,d) = (a + b,c + d)

We will make sure to explain this more clearly in a final version of the paper.

                                -------------

     - L718 and L719: Different kinds of data definitely have different
       shapes. I cannot understand why assigning a fixed probability will fit
       in all situations. Could you explain it a bit?

See the main response above the fold.

                                -------------

     - L866 - L868: The user provides two REs describing the source and view
       as input, which should already give some useful structure. For
       example, the top-level nonterminal may be described by several other
       non-terminals. This structure indeed suggests that a big problem can
       be divided into several subproblems.  Why do the user still need to
       manually "generate subproblems"? Is it avoidable?

We do not believe such subproblems are straightforwardly avoidable. While we can
attempt to match up subcomponents, this matching oftentimes will not work
because of, for example, issues where various equivalences must be applied (for
example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not break down. If we
had attempted to initially try breaking down the problem syntactically, we would
have wasted time. This is particularly an issue because bad specifications could
take a very long time, as there may be no good lenses between the attempted
proposed subcomponents, so there would potentially be a great deal of wasted
time in these attempts.

                                -------------

     - L900: Before introducing the benchmark suite, the authors had better
       provide the specification, regarding CPU, memory, etc., of their
       benchmark environment since an important measurement is the speed of
       the synthesis process.

All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of
1600 MHz DDR3 running macOS High Sierra.

                                -------------

     - L909 and L956: About the benchmark suite, how many of the benchmarks
       in fact only need bijective lenses? This definitely should be
       considered since (1) the benchmark suite is adapted from the "original
       Optician benchmarks, comprised of 39 bijective synthesis tasks", and
       (2) "Adjusting the termination parameter was needed in only 9 of the
       48 tasks". I suppose it is not, but still wonder, if for all of the 48
       - 39 = 9 non bijective tasks, we need to adjust their termination
       parameters?

35 of the benchmarks only need bijective lenses (4 of the original Optician
benchmarks had to be altered to be bijective because of projected data, in the
symmetric benchmark suite we removed such alterations.) Of the 9 benchmarks that
required the termination parameter, 4 of them required only bijective lenses.

                                -------------

     - L922: The authors may add a few words on how representative and
       complex the selected tasks is. For example, I got the data from the
       paper *Deep API Programmer: Learning to Program with APIs* that there
       is 238 real-world FlashFill benchmarks; then why the 8 data cleaning
       tasks here are more representative among the 238 tasks? The reviewers
       (and readers in the future) may want to know to which extent the job
       of writing lenses can be automatically done.

We chose our benchmarks from the FlashFill paper because the 238 tasks are,
unfortunately, not publicly available. In general, our benchmark suite goes
beyond the level of difficulty required for data cleaning. Many real-life
file format descriptors involve large numbers of disjunctions and nested
iterations, which FlashFill is either bad at (for disjunctions) or cannot
handle at all (nested iterations).

A more thorough comparison to FlashFill is present in the previous Synthesizing
Bijective Lenses paper.

                                -------------

     - L926 and L952: This is the most confusing point for me. The meaning of
       "correct lenses" here seems very subjective as the authors let the
       user determine whether the synthesised lens is desired. Can we make
       this process a bit automatic and objective?

       For instance, usually a natural and reasonable approach is to divide
       examples into a "training" set and a "validation" set; we use the
       training set to synthesise desired lenses and use the validation set
       to test if the synthesised lenses work well (although this does not
       prove its correctness). Even if the user is finally involved in
       determining whether the lens is a desired one, this process may reduce
       the burden of the user.

       (I also wonder how can the user verifies whether a complex lens of
       hundreds or thousands of lines code is the desired one by directly
       inspecting the lens program?)

We responded to this in detail above the fold.

                                -------------

     - L945 and L993 (Fig.6 and Fig.7): Since the time is only within 30
       seconds, why not make Figures whose x-axes are linear? With the
       "non-linear figures", I cannot tell that SS took 1.3 times longer than
       BS. In addition, since we only care about the average time of tasks
       (and the maximum time of a single task), there is no reason to use a
       line chart, whose shape is drastically affected by the order of
       tasks. A simple table is better, if the details of tasks (and their
       order) are not given.

We chose a logarithmic x axis to show the differences on the fast benchmarks, as
they otherwise would be imperceptable. However, if the impact on comparing the
slower benchmarks is too significant, we can change to a linear x axis.

Our graph does not include any specific order of tasks. At any given time, a
reader can see how many of the benchmarks are able to complete within that time
(for example, 41 benchmarks require complete in under a second in the full
synthesis algorithm). We can reword the description of these graphs to make this
more clear.

                                -------------

     - L1066: When discussing symmetric lenses, the authors need to give a
       reference to Symmetric Lenses by Martin Hofmann et al. (2011) since
       the definitions and notations are from there.

       L1079: The authors silently move to introduce edit lenses, although
       different from Edit Lenses by Martin Hofmann et al. (2012), a short
       introduction and a reference are needed. Moreover, the authors may
       need to explicitly mention that the edits in this section are
       "overwrites"; otherwise it is difficult to understand why the edits
       have the same type as the source and view (ignoring the sum types made
       by inl and inr.), as usually they are some "deltas".

We will update the lens section of the related work to be more thorough on the
related work that we do not make formal statements about, and include proper
citations.
found in the appendix.

                                -------------

     - L1111 and L1112: I cannot figure out where does `x` below the
       "horizontal line" come from

This is true for any arbitrary "x"

===========================================================================
Review #43C

    - The usability aspect is questionable: It seems that if the user has 
      defined complex regular expressions defining the pair of languages, then
      they can probably go ahead and just implement the lens without resorting 
      to examples.

As a demonstration about the difficulty of writing lenses, consider the
following bidirectional program

let single_author_convert =
  ins " au - "
    . lens_swap
        (NAME . del ",")
        (lens_swap WSP NAME)*
in

Even the relatively simple lens that permutes the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses comprised of many sub-lenses (though
for brevity we give here the more minimal example above).  Lenses provide
great power with their invertibility guarantees, but they come at the cost
of thinking about many fiddly details when writing the terms.  Worrying
about these fiddly details while ALSO thinking about unambiguity
restrictions is a very difficult task!

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not require these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between. Furthermore, the types provide resilience in the face of
future lens modifications, as the types ensure that the lens maps exactly
between strings of the provided formats.  Invalid inputs are not accepted
nor are the outputs ever invalid. Consequently, in a well-maintained
codebase, we are not requiring additional work. Instead, we are saving
programmers from having to writing the lens by leveraging work they should
already be performing.

                                -------------

    - More important: It is unclear to the general reader why the Expand
      procedure is needed — i.e., why do we need to rewrite types for
      GreedySynth.

We will make this more clear in future versions of the paper. Expand is
needed because GreedySynth alone cannot traverse regular expression
equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
(also with a disconect).  However, contuing search through equivalent REs
helps find the bijective lens.

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily above.  For a specific example:

Consider finding a lens of type:
("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "x", and
keep "c" mapped to "y".

Consider finding a lens of type:
("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "y", and
keep "c" mapped to "x".

We are happy to provide an example such as this one in the paper.

                                -------------

    - The paper resorts to a fixed 
      distribution that gives the same probability to each disjunct. This raises
      two concerns:

      1.  that this overfits to the hand-picked benchmarks and 
      2.  that the entropy calculation and stochastic REs are an overkill.

      Why not use a syntactic heuristic. That’s what DC is trying to do, but 
      perhaps a slightly more sophisticated syntactic heuristic could achieve 
      the same effect. So, while the use of entropy is cool, appealing, and new 
      in synthesis (to my knowledge), the paper does not convince me that it is 
      needed here.

We respond to these issues in the main reponse.

                                -------------

    - Section 6 or 7 should discuss examples that are outside the expressivity 
      of the lens language and/or take too much time. It was hard for me to 
      infer what the limitations of the approach are.

We agree that section 7 requires more, informal comparisons to other lenses, for
us to discuss comparisons between our framework and other types of lenses. In
general, for different types of lenses to work, we require additional work,
either in the synthesis algorithm (for different types of combinators like
matching lenses or delta lenses), or in both the synthesis algorithm and
specifications (like for edit lenses and update lenses).

                                -------------

    - My main concern is whether the use of stochastic REs is really useful (see
      above)

Stochastic REs, and more generally, a semantic cost metric is critical for
comparing between responses by GreedySynth. Without this semantic cost, the same
function would likely have different costs if found through different calls to
GreedySynth. Syntactic metrics, we feel, aim to approximate how good the
underlying result is: in this work we are able to find a semantic metric that we
can be found using syntactic information.

                                -------------

    - Are stochastic REs new? I didn't see any citations when they were 
      introduced.

Stochastic REs are not new. We cite them when we first mention them on page
2, but agree we should provide additional citations when they are formally
introduced. We do believe our syntactic technique for calculating their
entropy is new.