### 2018-09-20 papers/SymmetricLenses/response_popl19.md
Polishing response

BCP: The first section is too long -- 500 words max.

Thank you for your comments!

We begin with brief responses to the most significant concerns, followed by
an (optional) appendix addressing individual questions point by point.

                                -------------

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

We agree we did not go into enough detail on this, and we can add more in a
final version.

We do not have an automated way to measure the correctness of our lenses
(beyond checking that each synthesized lens does indeed transform the given
examples correctly). In this respect, our work is similar to other published
synthesis-from-examples research, such as Gulwani's work on FlashFill and
related projects.

To assess the behavior of our tool on our benchmarks, we did what
programmers usually do:

1.  We manually inspected the code.

    This was a nontrivial task, as generated lenses can be quite large, but
    it was easier than it might initially sound. Some of the lenses'
    structure should follow from the structure of the regular expressions
    describing source and target formats; this is relatively easy to
    validate. The more difficult part was checking the interactions of
    complex combinators like swaps, disconnects, and merges. Still these
    bits typically spanned only a few dozen lines of code, making our task
    easier. Certainly, checking the code is easier than creating it de novo.

2.  We ran a set of tests to validate the more complex lenses.

BCP: #2 needs a few more words

                                -------------

      - A number of related questions about our information-
        theoretic measure: Is it useful? Why not use a 
        (perhaps complex) syntactic metric? 
        Why does a fixed distribution work?  
        Will knowing the data distribution help?

BCP: Difficult, or impossible???

It is difficult (impossible) to prove that our information-theoretic measure
is better than all possible "syntactic" metrics, but our experiments show
that it performs well.  In particular, it outperforms simple syntactic
heuristics.  More important is the fact that our heuristic is based on a
simple, clear idea that is easy to communicate and implement.  This idea can
potentially be reused in a variety of related settings (and perhaps
_augmented_ with domain-specific heuristics).

BCP: I'd delete this: "By contrast, ad hoc syntactic metrics provide little
value beyond the single artifact in which they are used."

Moreover, our information-theoretic presentation has helped us understand
some of the ad hoc metrics used in the past -- [BCP: Whose?] those in which
people assign higher costs to "less bijective" combinators in simpler ways
(avoiding constants, for example).  These ad hoc syntactic metrics seem to
be approximating something like the information-theoretic measure we
work with.

A fixed distribution works because the core goal is to generate "more
bijective" transformations; the intuition is that information should be
preserved, when possible, when data is translated between formats.  A fixed,
uniform distribution avoids unfairly weighting one conjunct or disjunct over
another and unfairly throwing information away.  In the absence of
additional information, this kind of fairness seems to be the best we can
do.  [BCP: This is not very satisfying.  Seems like the question is exactly
"What kind of additional information might be useful and how could it be
communicated?"  Perhaps we could comment that this could be a direction for
future work.]

BCP: Oh, maybe the next paragraph is the discussion I was just asking for.
But I'm confused by the way it's introduced: Is this something we do now, or
something we could do?

On the other hand, knowing the distribution can help the system make a
choice about what to throw away: given a choice, it prefers to throw away
less data.  For example, let (x?.9) mean "x appears with probability .9 and
is "" otherwise and consider a conversion between these formats:

    (x?.999).(y?.001) <==> z

Here, x appears more frequently than y on the left.  Hence, the system will
prefer a lens that projects y and maps x to z over a lens that projects x
and maps y to z, as the latter tends to throw away much more information and
is "less bijective" in an information-theoretic sense.

===========================================================================
===========================================================================
End of main response.

Detailed responses to individual questions in the reviews follow...

===========================================================================
Review #43A

                                -------------

     - I have no significant complaints about this paper apart from Section 3 
       not making it clear that stochastic regular expressions are not a
       contribution of the authors (albeit it is mentioned in passing on p2).

Yes, we will make this clearer on p. 3 too.

                                -------------

     - 3. p5 l236,

          Shouldn't the second argument to `disconnect` be `company`?

Correct. However, in hindsight, we should change the example such that the
second type is "" (to be consistent with what we use in the full lens)

                                -------------

     - 4. p6, l288,

          If I understood correctly, this parituclar lens `disconnect(...)`
          picks the default values at random, but is still valid because it
          satisfies the required lens type. If so, please make it clearer.

This is correct -- we will clarify.

                                -------------

     - 9. p16, l777,

          Could you give concrete examples of situations where all sequences
          are involved in sequence lenses, but there are examples that could
          make other lenses more useful?

Within a DNF lens itself, all sequences are *required* to be involved with
another sequence lens: otherwise, when that side of the DNF SRE is provided,
what disjunct on the other side should that lens be mapped to? Of course,
there are other ways the disjunct can be mapped, but these must happen at a
different level of the synthesis algorithm. For example, the entirety of
that DNF regular expression (which itself could be deeply nested within
other DNF regular expression components) could be projected. Furthermore, a
specific sequence could be mapped to another, but all the components of that
sequence lens actually may be disconnects. However, all sequences *must* be
involved in a sequence lens in a well-typed DNF lens.

                                -------------

     - 10. p18, l842

           Could you explain in some more detail what criteria does `EXPAND` use
           to decide to open some closed expressions but not others?

At a high level, we process both regular expression trees, looking for
situations where a regular expression is a subcomponent of one format but
not the other, and expand such REs (this is where hash-consing provides its
greatest benefits). Also, we look for instances where a regular expression
may be present in one format, but is only contained within a "closed"
regular expression in the other, and we expand the closed RE containing the
missing RE in this case. We make this apply in a few more places by taking
into account how deeply nested within stars certain REs occur.

BCP: What is "closed"?
AFM: The internals of the Regular Expression are not exposed to GreedySynth,
they are the analogue of User Defined regular expressions in the previous paper,
but with this closed/open approach, we aren't dependent on where users place
variables (also making our integration into Boomerang significantly less hacky).

BCP: OK.  Can a short version of that make it into a parenthesis after
"closed" above?

Full details can be found in the prior work on Synthesizing Bijective Lenses
(page 16).

                                -------------

     - 11. p19, l887

           Surely the definition of `CONTINUE(pq)` should be using `max`? 
           Otherwise I don't see how the definition makes sense, because using 
           surely `min` will just pick 0 for t>=0, because the distance and the
           logarithm are positive values.

Right, this should be max!

                                -------------

     - 12. p21, l1022

           Which preference metric are the t=0, t=25, t=-25 benchmarks using?

These benchmarks are using the full "SS" mode. The only difference between these
benchmarks and the full mode is that termination parameters are ignored, and
are instead required to use the 0, 25, and -25 parameters.

                                -------------

      - Second, from what I remember from the literature on (ordinary) symmetric
        lenses, they do not lend themselves very well to analysis in terms of 
        spans of asymmetric lenses. With simple symmetric lenses, do you think
        the situation is better now that you have omitted the complement part?

It may be slightly better, but the details are tricky. From "Spans of Delta
Lenses," p1, equivalence relations are needed in both the span-presentation
and set-based presentation of symmetric lenses to address "'hidden' details
such as their *complements*." Simple symmetric lenses intentionally have no
such details. However, there seem to be other complexities related to lens
laws, requiring restrictions like very well-behavedness, as (from "Spans of
Lenses", p1) "in the category whose morphisms are lenses, what one might
expect to be the pullback of lenses need not satisfy the universal property
of a pullback." We do not believe our restriction helps obviate these
difficulties.

                                -------------

       - Thirdly, do the authors expect to be able to extend the proposed 
         synthesis algorithm to handle any of the many variants of updates-based
         lenses, such as edit lenses, delta lenses, and update lenses?

Yes, we do have some hope that the algorithm can be extended to handle such
variants.

Delta lenses would require additional work in understanding "shape"
constructs.  Currently our language only permits basic iteration, whereas
shapes allow for different types of mergings. Allowing this would require
changes to the AtomSynth algorithm, in that we would we would need to
discover the correct shapes of the data, and align them in the correct ways.

Both edit and update lenses would require new forms of specifications, since
both permit updates that are not merely pushing data through, but instead
are applying specific elements of an edit monoid.  (Also we would remove
complements from edit lenses).

===========================================================================
Review #43B

     - The main weakness is in the evaluation part. There are many 
       unclear/controversial parts regarding the evaluation. For example, the 
       lack of description of the benchmark environment; the way they test 
       whether the synthesised lens is correct; the description of the tasks to
       be synthesised. See detailed comments below. (This defect does not
       prevent it from being a good paper.)

We responded to the key points in this comment (the benchmark suite and
judging correctness) "above the fold" at the top of this response.

All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
GB of 1600 MHz DDR3 running macOS High Sierra.

                                -------------

     - The first subsection of Related Work did not give any reference to
       the related work except for Foster et al. 2007, which makes the
       paper unfriendly and hard to verify. The appendix is also in a
       messy.

We apologize for the lack of citations and the messy appendix. For the first
subsection of Related Work, we got carried away in making formal claims, and did
not include informal comparisons to other lens languages. We will make sure to
make such comparisons and citations in the next version of the paper. We will
similarly improve the appendix.

                                -------------

     - L11: The reference Martin Hofmann et al. (2015) does not help since it
       is not yet published and cannot be obtained from Internet. Martin
       Hofmann et al. (2011) may be the choice for this time.

Thank you, you are correct, we will change this citation.

                                -------------

     - L442: H(s) should be 0. Otherwise the entropy of the SRE in L452 is 2.

You are correct.

                                -------------

     - L484 and L485:
       I cannot see the reason to relax the restriction regarding unambiguous
       REs. What will happen if primitive lens combinators for unambiguous
       REs are composed with the ones for ambiguous REs? Since complex lenses
       are composed from (ambiguous and unambiguous) primitive lenses, in
       practice the restriction should be always there.

Within our synthesis algorithm, you are, in general, correct. However,
Boomerang exists beyond Optician, and permits lenses like id("a"|"a"). In
our extension to Boomerang, we do not remove this freedom.

                                -------------

     - L557 and below: Why some of the entropy is a single value while some
       of them seems ranging over an interval? If the interval is used to
       handle `merge_left`, then the entropy of all combinators should be an
       interval. Currently only `id`, `disconnect`, and `merge_left` yield
       intervals.

This was us trying to make our equations more readable, but we did not
sufficiently explain what our syntax meant. The calls to H(...) return pairs,
and when multiplied by constants, or added together, they act as follows:

  a x (b,c) = (a x b,a x c)
  (a,b) + (c,d) = (a + b,c + d)

We will make sure to explain this more clearly in a final version of the paper.

                                -------------

     - L718 and L719: Different kinds of data definitely have different
       shapes. I cannot understand why assigning a fixed probability will fit
       in all situations. Could you explain it a bit?

See the main response "above the fold" at the top of this response.

                                -------------

     - L866 - L868: The user provides two REs describing the source and view
       as input, which should already give some useful structure. For
       example, the top-level nonterminal may be described by several other
       non-terminals. This structure indeed suggests that a big problem can
       be divided into several subproblems.  Why do the user still need to
       manually "generate subproblems"? Is it avoidable?

We do not believe such subproblems are straightforwardly avoidable. While we
can attempt to match up subcomponents, this matching oftentimes will not
work because of, for example, issues where various equivalences must be
applied (for example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not
break down. If we had attempted to initially try breaking down the problem
syntactically, we would have wasted time. This is particularly an issue
because bad specifications could take a very long time, as there may be no
good lenses between the attempted proposed subcomponents, so there would
potentially be a great deal of wasted time in these attempts.

                                -------------

     - L900: Before introducing the benchmark suite, the authors had better
       provide the specification, regarding CPU, memory, etc., of their
       benchmark environment since an important measurement is the speed of
       the synthesis process.

All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
GB of 1600 MHz DDR3 running macOS High Sierra.

                                -------------

     - L909 and L956: About the benchmark suite, how many of the benchmarks
       in fact only need bijective lenses? This definitely should be
       considered since (1) the benchmark suite is adapted from the "original
       Optician benchmarks, comprised of 39 bijective synthesis tasks", and
       (2) "Adjusting the termination parameter was needed in only 9 of the
       48 tasks". I suppose it is not, but still wonder, if for all of the 48
       - 39 = 9 non bijective tasks, we need to adjust their termination
       parameters?

35 of the benchmarks only need bijective lenses (4 of the original Optician
benchmarks had to be altered to be bijective because of projected data, in
the symmetric benchmark suite we removed such alterations.) Of the 9
benchmarks that required the termination parameter, 4 of them required only
bijective lenses.

                                -------------

     - L922: The authors may add a few words on how representative and
       complex the selected tasks is. For example, I got the data from the
       paper *Deep API Programmer: Learning to Program with APIs* that there
       is 238 real-world FlashFill benchmarks; then why the 8 data cleaning
       tasks here are more representative among the 238 tasks? The reviewers
       (and readers in the future) may want to know to which extent the job
       of writing lenses can be automatically done.

We chose our benchmarks from the FlashFill paper because the 238 tasks are,
unfortunately, not publicly available. In general, our benchmark suite goes
beyond the level of difficulty required for data cleaning. Many real-life
file format descriptors involve large numbers of disjunctions and nested
iterations, which FlashFill is either bad at (for disjunctions) or cannot
handle at all (nested iterations).

A more thorough comparison to FlashFill can be found in the previous
Synthesizing Bijective Lenses paper.

                                -------------

     - L926 and L952: This is the most confusing point for me. The meaning of
       "correct lenses" here seems very subjective as the authors let the
       user determine whether the synthesised lens is desired. Can we make
       this process a bit automatic and objective?

       For instance, usually a natural and reasonable approach is to divide
       examples into a "training" set and a "validation" set; we use the
       training set to synthesise desired lenses and use the validation set
       to test if the synthesised lenses work well (although this does not
       prove its correctness). Even if the user is finally involved in
       determining whether the lens is a desired one, this process may reduce
       the burden of the user.

       (I also wonder how can the user verifies whether a complex lens of
       hundreds or thousands of lines code is the desired one by directly
       inspecting the lens program?)

See our detailed response "above the fold."

                                -------------

     - L945 and L993 (Fig.6 and Fig.7): Since the time is only within 30
       seconds, why not make Figures whose x-axes are linear? With the
       "non-linear figures", I cannot tell that SS took 1.3 times longer than
       BS. In addition, since we only care about the average time of tasks
       (and the maximum time of a single task), there is no reason to use a
       line chart, whose shape is drastically affected by the order of
       tasks. A simple table is better, if the details of tasks (and their
       order) are not given.

We chose a logarithmic x axis to show the differences on the fast
benchmarks, as they otherwise would be imperceptable. However, if the impact
on comparing the slower benchmarks is too significant, we can change to a
linear x axis.

Our graph does not include any specific order of tasks. At any given time, a
reader can see how many of the benchmarks are able to complete within that
time (for example, 41 benchmarks require complete in under a second in the
full synthesis algorithm). We can reword the description of these graphs to
make this more clear.

                                -------------

     - L1066: When discussing symmetric lenses, the authors need to give a
       reference to Symmetric Lenses by Martin Hofmann et al. (2011) since
       the definitions and notations are from there.

       L1079: The authors silently move to introduce edit lenses, although
       different from Edit Lenses by Martin Hofmann et al. (2012), a short
       introduction and a reference are needed. Moreover, the authors may
       need to explicitly mention that the edits in this section are
       "overwrites"; otherwise it is difficult to understand why the edits
       have the same type as the source and view (ignoring the sum types made
       by inl and inr.), as usually they are some "deltas".

We will update the lens section of the related work to be more thorough on
the related work that we do not make formal statements about, and include
proper citations.

                                -------------

     - L1111 and L1112: I cannot figure out where does `x` below the
       "horizontal line" come from

This is true for any arbitrary x (as that x gets "overwritten" by the user
provided x')

===========================================================================
Review #43C

    - The usability aspect is questionable: It seems that if the user has 
      defined complex regular expressions defining the pair of languages, then
      they can probably go ahead and just implement the lens without resorting 
      to examples.

As a demonstration about the difficulty of writing lenses, consider the
following bidirectional program

     let single_author_convert =
       ins " au - "
         . lens_swap
             (NAME . del ",")
             (lens_swap WSP NAME)*
     in ...

Even the relatively simple lens that permutes the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses built from many sub-lenses.  Lenses
provide great power with their invertibility guarantees, but they come at
the cost of thinking about fiddly details when writing the terms.  Worrying
about these fiddly details while ALSO thinking about unambiguity
restrictions is a difficult task!

Engineering best practices suggest that it's a good idea to annotate lenses
with their types.  While Boomerang does not require these annotations (it
can infer the types from the terms), types serve as documentation for future
programmers to understand what formats the lens maps between. Furthermore,
the types provide resilience in the face of future lens modifications,
ensuring that the lens maps exactly between strings of the provided formats.

                                -------------

    - More important: It is unclear to the general reader why the Expand
      procedure is needed — i.e., why do we need to rewrite types for
      GreedySynth.

Expand is needed because GreedySynth alone cannot traverse regular
expression equivalences. If given the two regular expressions "" | "a"+, and
"a"\*, GreedySynth alone would merge "" to "a"\* (with a disconect), and
merge "a"+ (also with a disconect).  However, contuing search through
equivalent REs helps find the bijective lens.

                                -------------

    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
      point that if the user knew the distribution (e.g., if they automatically
      infer the parameters from data), then they can supply it. But will that
      really help the search? Essentially, I wish there were examples (even 
      synthetic ones) where knowing the distribution of data can clearly 
      influence the synthesis procedure.

We respond to this primarily in the main response "above the fold" at the
top of this response. For a specific example:

Consider finding a lens of type:

    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "x", and
keep "c" mapped to "y".

Consider finding a lens of type:

    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"

In this situation, the lowest cost lens would merge "a" and "b" into "y", and
keep "c" mapped to "x".

We can provide an example like this in the paper.

                                -------------

    - The paper resorts to a fixed 
      distribution that gives the same probability to each disjunct. This raises
      two concerns:

      1.  that this overfits to the hand-picked benchmarks and 
      2.  that the entropy calculation and stochastic REs are an overkill.

      Why not use a syntactic heuristic. That’s what DC is trying to do, but 
      perhaps a slightly more sophisticated syntactic heuristic could achieve 
      the same effect. So, while the use of entropy is cool, appealing, and new 
      in synthesis (to my knowledge), the paper does not convince me that it is 
      needed here.

See the main reponse.

                                -------------

    - Section 6 or 7 should discuss examples that are outside the expressivity 
      of the lens language and/or take too much time. It was hard for me to 
      infer what the limitations of the approach are.

We agree that section 7 requires more, informal comparisons to other lenses,
for us to discuss comparisons between our framework and other types of
lenses. In general, for different types of lenses to work, we require
additional work, either in the synthesis algorithm (for synthesizing
different types of combinators like matching lenses or delta lenses), or in
both the synthesis algorithm and specifications (like for edit lenses and
update lenses).

                                -------------

    - My main concern is whether the use of stochastic REs is really useful (see
      above)

Stochastic REs, and more generally, a semantic cost metric are critical for
comparing between responses by GreedySynth. Without this semantic cost, the
same function would likely have different costs if found through different
calls to GreedySynth. Syntactic metrics, we feel, aim to approximate how
good the underlying result is: in this work we are able to find a semantic
metric that we can be found using syntactic information.

                                -------------

    - Are stochastic REs new? I didn't see any citations when they were 
      introduced.

Stochastic REs are not new. We cite them when we first mention them on page
2, but agree we should provide additional citations when they are formally
introduced. We do believe our syntactic technique for calculating their
entropy is novel.