### 2019-02-12 papers/BXProposal/main.tex
wordsmithing and updated abstract

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{onecolpceurws}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\title{Expanding the Power of Lens Synthesis \\ (talk proposal)}

% I've never done a talk before, should this be everybody or just me
\author{
Anders Miltner \\ Princeton University \\ \ 
% Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
\and
Solomon Maina \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
\and
Kathleen Fisher \\ Tufts University
% Medford, MA 02155 \\ kfisher@eecs.tufts.edu
\and
Benjamin C. Pierce \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
\and
David Walker \\ Princeton University
% Princeton, NJ 08540 \\ dpw@cs.princeton.edu
\and
Steve Zdancewic \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
}

\institution{}

\begin{document}
\maketitle

\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.4,0,0.6}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{vlightyellow}{RGB}{255, 255, 200}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{vlightgreen}{RGB}{190, 255, 230}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const,skip},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}

\begin{abstract}
  {\color{dkblue} \emph{AFM: I will remove this from the submission, but am including it in the
    document for people to review. It is due today (February 12)}} Recent work
  has shown how to synthesize a class of string bijections from regular
  expression specifications. These bijective string transformations can be
  useful for synchronizing simple data formats, but many interesting formats
  require non-bijective lenses. In this talk, we show how to expand the class of
  synthesized transformations beyond bijections to quotient lenses and a large
  subset of full symmetric lenses (including all the standard asymmetric lenses)
  through lightweight user annotations that help identify a ``bijective core.''
\end{abstract}

\section{Synthesizing Bijective Transformations}

Bidirectional languages can be difficult to program in, proposing unfamiliar
programming paradigms, novel syntax, and subtle typing constraints.
Fortunately, their restricted expressive power, together with the
fine-grained type systems needed to guarantee that well-types programs
satisfy strong round-tripping properites, make them a perfect target for
program synthesis.

The {\em Optician} tool is an extension to Boomerang~\cite{boomerang}---a
DSL for writing lenses~\cite{Focal2005-long2} on strings---that synthesizes
{\em bijective} Boomerang lenses when provided with a pair of types (regular
expressions) and a small number of examples (often zero).

\section{Building Complex Transformations With Bijective Cores}

Bijections are an important class of bidirectional transformations, but many
useful transformations are not completely bijective. In particular, Boomerang
expressions are not necessarily bijections -- they encode lenses and
various lens extensions like quotient lenses~\cite{quotientlenses}. Lenses are
bidirectional transformations that may lose information when going from one side
to the other, while obeying certain roundtripping properties. Quotient lenses
obey similar properties to lenses, but also permit treating certain pieces of
information as irrelevant -- normalizing them to a canonical representative.

While we wish to make creating these lenses easier, Optician's synthesis
algorithm only works on bijective lenses. Fortunately, many lenses and quotient
lenses seem to have a ``bijective core'', a component that merely changes the
shape of data, but does not remove any data. This bijective component can then
be surrounded by projection and quotient components to form a full-fledged
quotient lens.

Surrounding bijections in this way can be already be done with Optician
and Boomerang, but doing so requires repeated code. Consider the
converting an index of musical composers (with names, birth and death
dates, and nationalities) to a simple index of musical composers (with just
names and nationalities and irrelevant whitespace). In this example, ``.''
represents concatenation and ``\lstinline{*}'' represents the Kleene star.

\begin{lstlisting}
  let full_composers : regex = (name . dates . nationality)*
  let composers_nodates : regex = (nationality . " "* . name)*
\end{lstlisting}

In this example, the bijective core is the bijection that transmits nationality and
names (while swapping them), the projection component removes \lstinline{dates}
when operating from left-to-right and recovers it when operating from right-to-left,
and the quotient component normalizes the whitespace to the empty string.

We can build a projection component for \lstinline{full_composers} using
existing Boomerang combinators.
%
\begin{lstlisting}
  let full_projection = iterate (name . (del dates) . nationality)
\end{lstlisting}
%
Here, \lstinline{del} deletes data when going in one direction, and restores it
when going backwards.

We can also build a normalization component for \lstinline{dates} using existing
Boomerang combinators.
%
\begin{lstlisting}
  let ws_norm = normalize (del " "*)
  let nodates_norm = iterate (nationality . (rquot ws_norm) . name)
\end{lstlisting}
%
Here, \lstinline{normalize} turns a lens into a ``canonizer'' (in effect, a
normalization function). In this example, \lstinline{normalize} canonizes an
arbitrarily long list of whitespace to the empty string. Then, the canonizer is
hooked into a lens with \lstinline{rquot}, normalizing strings when the lens is
applied right-to-left, and performing the identity when applied from left-to-right.

With these normalization and projection components defined, the bijective core
can be synthesized (using \lstinline{synth $R$ <=> $S$}), and the full lens can be
provided by composing all the subcomponents (using ``;'').
\begin{lstlisting}
  let bij_core = synth (targetType full_projection) <=> (sourceType nodates_norm)
  let composers : full_composers <=> composers_nodates = full_projections ; bij_core ; nodates_norm
\end{lstlisting}

Note how in both the projection and quotient component, the bijective aspects
are repeated with no real alteration -- they are merely required scaffolding
for expressing the projections and normalizations.

\section{Projection and Quotient Annotations}

Instead of manually writing explicit canonization and projection functions, we
consider an alternative approach: by annotating our regular expressions with
canonization and projection annotations, we can write our data formats and find
the bijective core in a single step. This approach has already been employed in
synthesizing bidirectional transformations that are bijections (up to
quotients)~\cite{maina+:quotient-synthesis}, but here we extend it with projections as well.

Consider the left-hand format of \lstinline{full_composers}. We expand our
regular expression types with the new combinator \lstinline{skip}. Annotating a
regular expression to be skipped does exactly what \lstinline{full_projection}
does in the previous code -- it sends the format, after removing the projected
portions, to the core synthesis engine, then composes the generated lens with a
lens that deletes the regular expression from the data format.
%
\begin{lstlisting}
  let full_composers' : annotated_regex = (name . (skip dates) . nationality)*
\end{lstlisting}
%

Consider the right-hand format of \lstinline{composers_nodates}. We expand our
regular expression types with a number of new combinators (to address different
types of quotients), one of which is
\lstinline{squash}~\cite{maina+:quotient-synthesis}. Annotating a regular
expression to be squashed does exactly what \lstinline{nodates_norm} does in the
previous code -- it sends the normalized format to the core synthesis engine,
then composes the generated lens with a lens that normalizes the whitespace to a
representative (in the example, the empty string).
%
\begin{lstlisting}
  let composers_nodates' : annotated_regex = (nationality . (squash " "*) . name)*
\end{lstlisting}
%

Now, with annotated regular expressions, we can synthesize the composers lens much
more compactly.
%
\begin{lstlisting}
  let composers'  = synth full_composers' <=> composers_nodates'
\end{lstlisting}
%
Synthesizing with these annotated regular expressions performs the same steps
that we previously did manually: it breaks the lens into concatenation and
normalization subcomponents, sends the bijective core to the existing synthesis
algorithm, and composes the generated subcomponents with the synthesized
bijective core.

While we have shown how these annotations can work for quotient lenses, we could
even imagine getting them to work for symmetric lenses. If we viewed symmetric
lenses as also having a bijective core (with projections on both sides instead
of just one), providing \lstinline{skip} annotations on the right could even
make synthesizing symmetric lenses easier. Of course, this would require
symmetric lenses in Boomerang for the generated lenses to be well-typed.

Furthermore, we think this approach could scale to other lens extensions, like
matching lenses and dictionary lenses, and other types of bidirectional
constructs beyond merely lenses. All that is needed for this approach is a
user-provided a reduction to a ``bijective core,'' and a means to synthesize
that core. If a bijective synthesis algorithm is discovered for
(for example) trees or graphs, then annotations can be used to synthesize more
complex bidirectional programs for those structures as well.

\bibliographystyle{alpha} 
\bibliography{local}
%inline the .bbl file directly for mailing to authors.

\end{document}