### 2019-04-16 papers/SymmetricLenses/rebuttal_icfp19.txt
rebuttal

Thank you for your feedback. We respond the major questions and concerns, then
provide an (optional) point-by-point response to individual concerns. We are
happy to include any of the information presented in this rebuttal in the final
version of the paper.

    - Why is writing two regexes easier for the user than writing a lens? Do
      input formats often come equipped with regexes already? Can the
      synthesizer use such pre-existing regexes out of the box, or do they have
      to be "massaged" by the user?

While we require users to write regular expressions, programming regular
expressions is easier than programming lenses:

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while a lens deals with two
  at the same time.

* In the (common) case where the two formats have Kleene stars in different
  places, manual lens programming requires mentally transforming both to a
  common "aligned" form.  Our synthesis algorithm does this alignment
  automatically.  This alignment can be quite challenging for programmers.

As an explicit example, consider the following function:

        let last_head = NAME . "," . (WSP . NAME)*
        let last_tail = (NAME . WSP)* NAME

        let name_conversion : last_head <=> last_tail =
            lens_swap
                (NAME . del ",")
                (lens_swap WSP NAME)*

While the name_conversion expression may have more AST nodes than the two
regular expressions, we find it much harder to write.

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not require these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, the types provide resilience in the face of future
lens modifications, as the types ensure that the lens maps exactly
between strings of the provided formats. Invalid inputs are not accepted
nor are the outputs ever invalid.  Consequently, we are not requiring
additional work. Instead, we are saving programmers from having to 
writing the lens by leveraging work they should already be performing.

In our experiments, we manually programmed the regular expressions. While
writing these regular expressions, we did not massage them to be amenable to
synthesis. In general, the synthesizer is able to handle regular expressions
expressed in multiple different ways, due to the traversal of regular expression
equivalences. However, if the regular expressions require a transformation that
was not a star-semiring transformation, such a massaging would be required.

    - What is the relationship between our information-theoretic ranking, and
      prior work like Unsupervised Learning by Program Synthesis

(afm: this can potentially go after the fold, as it was only mentioned in
passing, not as a critical question of the reviewer)

Unsupervised Learning by Program Synthesis (ULPS) expands upon previous research
in unsupervised machine learning. Unsupervised learning attempts to "induce good
latent representations of a data set" [Ellis et al. 2015]. Given a set of data
points, or "outputs," ULPS attempts to find good representations of these
outputs as a combination of finding a single function, and a sequence of
"inputs" that correspond to the outputs. The latent representation is a
combination of the corresponding input, and the transformation from inputs to
outputs.

Our work is similar to this work in that we both use information-theoretic
principles to find the desired programs, but the use of information theory in
each is different. In ULPS, minimizing the information theoretic cost metric is
the only goal. In this metric, the desire is to encode the output data in as few
bits as possible, so the example data is tightly coupled with the problem. In
our work, we use information theory to formalize the notion of "more bijective,"
providing a firm theoretical foundation for previous ad-hoc ranking functions.
In our case, the examples serve merely as a specification, but the inclusion or
exclusion of such examples does not impact "how bijective" certain functions
are. We do not consider this a weakness, but an inherent aspect to our
formalization. Furthermore, we would not consider it a good idea to assume that
we should try to minimize the data recovery costs on the examples more than on
other pieces of data -- non-bijective input/output pairs are often useful.

    - The authors do not consider lens compositions in their synthesis. Are
      there any justifications not to consider them? In existing results that
      synthesize bijective (quotient) lenses, one of the reasons not to consider
      lens composition in their synthesis is that ruling out them does not
      change the expressive power? Can you say a similar argument for this case?

We do not synthesize lens compositions for the same reasons as previous work --
such compositions are quite difficult to generate. You are correct that we do
not prove that such a restriction does not limit the expressive power. However,
we do not consider the symmetric DNF lens language a core contribution
(symmetric DNF lenses are not even defined outside of the appendix) and do not
believe such a proof would add much to the paper. However, we do provide
experimental evidence that the composition operator is not needed in practice;
we are able to synthesize all our benchmarks.

    - There is no explanation of the selection of the benchmarks. What are their
      characteristics and why those exercise interesting aspects of the
      algorithm.

We describe the selection of the benchmark suite in section 6.1. 29 of the
benchmarks are taken from Augeas, a bidirectional tool for editing Linux
configuration files. These examples were quite large, and involved some very
complex transformations, though these benchmarks were typically bijective. These
demonstrated how our tool performed at large, complex transformations. The 8
benchmarks from Flash Fill provided a range of transformations, none of which
were not bijective. These demonstrated how our tool performed on tasks that lost
large amounts of data, and sometimes required extended transformations through
Expand. The remaining 11 benchmarks tested how our tool performed at the type of
transformations we are familiar with, from transformations that would be useful
in day-to-day software engineering or classic transformations from the
bidirectional programming literature.

    - There is no attempt to explain what are the sources of worse performance
      for SS.

The primary impact on speed is the use of GreedySynth, instead of the bijective
RigidSynth. For GreedySynth, lenses are found between regular expressions that,
ultimately, do not show up in the final lens. RigidSynth does not incur this
performance penalty -- it only synthesizes lenses between regular expressions
for which there exists a bijective lens between them.

=================================================================================
Reviewer A
=================================================================================

    - Moreover, on top of the regexes, the user needs to provide examples and
      relevance annotations, and the only way to do so seems to be through an
      iterative process (by inspecting the synthesized lens and adding more
      examples / annotations when needed). Such iterative approach would be
      challenging for a user who is not sufficiently familiar with lenses to
      write one themselves.

We described the iterative process assuming a lazy user who didn't want to
provide more specifications than were necessary. However, such an iterative
process is not required, and if the user is confident that certain information
is "required" or may safely "skipped," they can safely add this information --
it will not slow down synthesis. Similarly, if the user already has a benchmark
suite, they may safely include such examples.

    - To add to the previous point: the evaluation could do a better job
      quantifying these limitations. For example, it would be good to see for
      each benchmark the size of annotations required, as well as how many
      "existing lenses" were used in compositional synthesis and where they came
      from (some standard library of lenses? or did the authors manually
      decompose each problem into subproblems? if so, then how many subproblems
      were there?)

There were a total of 12 uses of "require" in the benchmark suite. There were a
total of 4 uses of "skip" in the benchmark suite. In total 73 existing lenses
were used in compositional synthesis (about 1.5 per benchmark on average).
Typically, we manually decomposed these lenses, though some used the to_upper
function. We are happy to include a figure detailing the subtasks specified
during compositional synthesis in the final version of the paper (similar to the
one on p24 of Synthesizing Bijective Lenses).

    - Finally, the evaluation could also do a better job confirming that there's
      indeed a benefit to synthesizing a lens, rather than using a tool like
      FlashFill to synthesize the two functions independently. The related work
      section mentions that FlashFill cannot handle nested loops. However,
      FlashFill has evolved a lot since its original publication in 2011, so it
      would be interesting to see an empirical comparison.

We included an extended comparison to FlashFill in previous work, and did not
include such a comparison for lack of space. We are happy to make comparison to
FlashFill cited more explicitly.

    - The introduction mentions that stochastic regular expressions guide the
      search process. I would like to see it better explained in the main text
      of the paper: do they actually guide the search or merely rank the
      results?

Due to the Greedy nature of GreedySynth, the stochastic regular expressions
guide the search (but the stochastic annotations only guide the search insofar
as the lens costs guide the search).

    - L 284: please explain why this rule is needed for the lens in Fig. 3 to be
      well-typed.

Without the type equivalence rule, the lens would be well-typed between
emp_salaries and header . ("" | ("\n" . emp_ins) . ("\n" . emp_ins)∗). To be
well-typed between emp_salaries and emp_insurance, the right-hand type needs to
be transformed into an equivalent form.

    - Section 5.6: expansion inference seems to be a crucial part of the
      algorithm but is described very briefly. I would like to see more detail.

Expansion inference is the same as is performed in bijective lens inference, and
we did not include it for space. We will gladly make such a citation more
explicit.

    - Can you explain why you need the lower bound [for intervals]?

While we do not need it in the algorithm, we include it to make formal claims
about bounding the cost. We describe the need for intervals in 4.1.